---
title: "Dataframe outputs"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    toc_depth: 3
    warning: FALSE
    message: FALSE
vignette: >
  %\VignetteIndexEntry{Dataframe outputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# show me all columns
options(
  tibble.width = Inf,
  pillar.bold = TRUE,
  pillar.neg = TRUE,
  pillar.subtle_num = TRUE,
  pillar.min_chars = Inf
)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  out.width = "100%",
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The returned dataframe will contain the following columns (the exact columns
will depend on the test and the statistical approach):

  - *statistic*: the numeric value of a statistic.

  - *df*: the numeric value of a parameter being modeled (often degrees
    of freedom for the test); note that if `no.parameters = 0L` (e.g., for
    non-parametric tests), this column will be irrelevant.

  - *df.error*, *df*: relevant only if the statistic in question has
    two degrees of freedom (e.g., anova).

  - *p.value*: the two-sided *p*-value associated with the observed statistic.
  
  - *method*: the details of the statistical test carried out.

  - *estimate*: estimated value of the effect size.

  - *conf.low*: lower bound for effect size estimate.

  - *conf.high*: upper bound for effect size estimate.
  
  - *conf.level*: width of the confidence interval.
  
  - *effectsize*: the details of the effect size.

For details about the tests and types of effect size estimates they correspond
to, see:
<https://indrajeetpatil.github.io/statsExpressions/articles/stats_details.html>

# One-sample tests

```{r onesample}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "parametric",
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "nonparametric",
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "robust",
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "bayes",
  bf.prior = 0.8,
)
```

# Two-sample tests

## within-subjects design

```{r twosample_w}
# setup
library(statsExpressions)

# data
df <- dplyr::filter(bugs_long, condition %in% c("LDLF", "LDHF"))

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "p",
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "bayes"
)
```

## between-subjects design

```{r twosample_b}
# setup
set.seed(123)
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
# unequal variance
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "p"
)

# for reproducibility
set.seed(123)
# equal variance
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  var.equal = TRUE,
  type = "p"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "bayes"
)
```

# One-way ANOVAs

## within-subjects design

```{r anova_w}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

if (require("afex", quietly = TRUE)) {
  # for reproducibility
  set.seed(123)
  oneway_anova(
    data = bugs_long,
    x = condition,
    y = desire,
    paired = TRUE,
    subject.id = subject,
    type = "p"
  )
}

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

if (utils::packageVersion("BayesFactor") >= package_version("0.9.12-4.3")) {
  # for reproducibility
  set.seed(123)
  oneway_anova(
    data = bugs_long,
    x = condition,
    y = desire,
    paired = TRUE,
    subject.id = subject,
    type = "bayes"
  )
}
```

## between-subjects design

```{r anova_b}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
# unequal variance
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "p"
)

# for reproducibility
set.seed(123)
# equal variance
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  var.equal = TRUE,
  type = "p"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "bayes"
)
```

# Contingency table analyses

```{r conttabs}
# for reproducibility
set.seed(123)
library(statsExpressions)

# create data structure
paired_data <-
  structure(
    list(
      response_before =
        structure(
          c(1L, 2L, 1L, 2L),
          .Label = c("no", "yes"),
          class = "factor"
        ),
      response_after = structure(
        c(1L, 1L, 2L, 2L),
        .Label = c("no", "yes"),
        class = "factor"
      ),
      Freq = c(65L, 25L, 5L, 5L)
    ),
    class = "data.frame",
    row.names = c(NA, -4L)
  )

# ------------------------ non-Bayesian -----------------------------

# association tests

# unpaired
set.seed(123)
contingency_table(
  data = mtcars,
  x = am,
  y = cyl,
  paired = FALSE,
)

# paired
set.seed(123)
contingency_table(
  data = paired_data,
  x = response_before,
  y = response_after,
  paired = TRUE,
  counts = "Freq"
)

# goodness-of-fit tests

# with counts
set.seed(123)
contingency_table(
  data = as.data.frame(HairEyeColor),
  x = Eye,
  counts = Freq,
)

# ------------------------ Bayesian -----------------------------

# association test
contingency_table(
  data = mtcars,
  x = am,
  y = cyl,
  paired = FALSE,
  type = "bayes"
)

# goodness-of-fit test
contingency_table(
  data = as.data.frame(HairEyeColor),
  x = Eye,
  counts = Freq,
  ratio = c(0.2, 0.2, 0.3, 0.3),
  type = "bayes"
)
```

# Correlation analyses

```{r corr}
# for reproducibility
set.seed(123)
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "parametric"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "nonparametric"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "robust"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "bayes"
)
```

# Meta-analysis

```{r meta}
# for reproducibility
set.seed(123)
library(statsExpressions)
library(metaplus)

# renaming to what `statsExpressions` expects
df <- dplyr::rename(mag, estimate = yi, std.error = sei)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
meta_analysis(data = df, type = "parametric")

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
meta_analysis(data = df, type = "robust")

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
meta_analysis(data = df, type = "bayes")
```

# Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: <https://github.com/IndrajeetPatil/ggstatsplot/issues>

# Session Information

For details, see- <https://indrajeetpatil.github.io/ggstatsplot/articles/web_only/session_info.html>
