---
title: "Dataframe outputs"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    toc_depth: 3
    warning: FALSE
    message: FALSE
vignette: >
  %\VignetteIndexEntry{Dataframe outputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# show me all columns
options(
  tibble.width = Inf,
  pillar.bold = TRUE,
  pillar.neg = TRUE,
  pillar.subtle_num = TRUE,
  pillar.min_chars = Inf
)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  out.width = "100%",
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The `statsExpressions` package has two key aims: to provide a consistent syntax
to do statistical analysis with tidy data, and to provide statistical
expressions (i.e., a pre-formatted in-text statistical result) for plotting
functions. Currently, it supports common types of statistical approaches and
tests: parametric, nonparametric, robust, and Bayesian *t*-test, one-way ANOVA,
correlation analyses, contingency table analyses, and meta-analyses. The
functions are pipe-friendly and compatible with tidy data.

Statistical packages exhibit substantial diversity in terms of their syntax and
expected input type. This can make it difficult to switch from one statistical
approach to another. For example, some functions expect vectors as inputs, while
others expect dataframes. Depending on whether it is a repeated measures design
or not, different functions might expect data to be in wide or long format. Some
functions can internally omit missing values, while other functions error in
their presence. Furthermore, if someone wishes to utilize the objects returned
by these packages downstream in their workflow, this is not straightforward
either because even functions from the same package can return a list, a matrix,
an array, a dataframe, etc., depending on the function.

This is where `statsExpressions` comes in: It can be thought of as a unified
portal through which most of the functionality in these underlying packages can
be accessed, with a simpler interface and no requirement to change data format.

# Dataframes

The dataframe will contain the following columns (the exact columns will depend
on the test and the statistical approach):

  - `statistic`: the numeric value of a statistic.

  - `df`: the numeric value of a parameter being modeled (often degrees
    of freedom for the test); note that if `no.parameters = 0L` (e.g., for
    non-parametric tests), this column will be irrelevant.

  - `df.error` and `df`: relevant only if the statistic in question has
    two degrees of freedom (e.g., anova).

  - `p.value`: the two-sided *p*-value associated with the observed statistic.
  
  - `method`: the details of the statistical test carried out.

  - `estimate`: estimated value of the effect size.

  - `conf.low`: lower bound for the effect size estimate.

  - `conf.high`: upper bound for the effect size estimate.
  
  - `conf.level`: width of the confidence interval.
  
  - `effectsize`: the type of the effect size.

For details about the tests and types of effect size estimates they correspond
to, see:
<https://indrajeetpatil.github.io/statsExpressions/articles/stats_details.html>

# One-sample tests

```{r onesample}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "parametric",
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "nonparametric",
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "robust",
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
one_sample_test(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "bayes",
  bf.prior = 0.8,
)
```

# Two-sample tests

## within-subjects design

```{r twosample_w}
# setup
library(statsExpressions)

# data
df <- dplyr::filter(bugs_long, condition %in% c("LDLF", "LDHF"))

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "p",
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "bayes"
)
```

## between-subjects design

```{r twosample_b}
# setup
set.seed(123)
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
# unequal variance
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "p"
)

# for reproducibility
set.seed(123)
# equal variance
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  var.equal = TRUE,
  type = "p"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
two_sample_test(
  data = ToothGrowth,
  x = supp,
  y = len,
  type = "bayes"
)
```

# One-way ANOVAs

## within-subjects design

```{r anova_w}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

if (require("afex", quietly = TRUE)) {
  # for reproducibility
  set.seed(123)
  oneway_anova(
    data = bugs_long,
    x = condition,
    y = desire,
    paired = TRUE,
    subject.id = subject,
    type = "p"
  )
}

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

if (utils::packageVersion("BayesFactor") >= package_version("0.9.12-4.3")) {
  # for reproducibility
  set.seed(123)
  oneway_anova(
    data = bugs_long,
    x = condition,
    y = desire,
    paired = TRUE,
    subject.id = subject,
    type = "bayes"
  )
}
```

## between-subjects design

```{r anova_b}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
# unequal variance
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "p"
)

# for reproducibility
set.seed(123)
# equal variance
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  var.equal = TRUE,
  type = "p"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "np"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "r"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
oneway_anova(
  data = iris,
  x = Species,
  y = Sepal.Length,
  type = "bayes"
)
```

# Contingency table analyses

```{r conttabs}
# for reproducibility
set.seed(123)
library(statsExpressions)

# create data structure
paired_data <-
  structure(
    list(
      response_before =
        structure(
          c(1L, 2L, 1L, 2L),
          .Label = c("no", "yes"),
          class = "factor"
        ),
      response_after = structure(
        c(1L, 1L, 2L, 2L),
        .Label = c("no", "yes"),
        class = "factor"
      ),
      Freq = c(65L, 25L, 5L, 5L)
    ),
    class = "data.frame",
    row.names = c(NA, -4L)
  )

# ------------------------ non-Bayesian -----------------------------

# association tests

# unpaired
set.seed(123)
contingency_table(
  data = mtcars,
  x = am,
  y = cyl,
  paired = FALSE,
)

# paired
set.seed(123)
contingency_table(
  data = paired_data,
  x = response_before,
  y = response_after,
  paired = TRUE,
  counts = "Freq"
)

# goodness-of-fit tests

# with counts
set.seed(123)
contingency_table(
  data = as.data.frame(HairEyeColor),
  x = Eye,
  counts = Freq,
)

# ------------------------ Bayesian -----------------------------

# association test
contingency_table(
  data = mtcars,
  x = am,
  y = cyl,
  paired = FALSE,
  type = "bayes"
)

# goodness-of-fit test
contingency_table(
  data = as.data.frame(HairEyeColor),
  x = Eye,
  counts = Freq,
  ratio = c(0.2, 0.2, 0.3, 0.3),
  type = "bayes"
)
```

# Correlation analyses

```{r corr}
# for reproducibility
set.seed(123)
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "parametric"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "nonparametric"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "robust"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "bayes"
)
```

# Meta-analysis

```{r meta}
# for reproducibility
set.seed(123)
library(statsExpressions)
library(metaplus)

# renaming to what `statsExpressions` expects
df <- dplyr::rename(mag, estimate = yi, std.error = sei)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
meta_analysis(data = df, type = "parametric")

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
meta_analysis(data = df, type = "robust")

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
meta_analysis(data = df, type = "bayes")
```

# Centrality description

```{r centrality}
# ----------------------- parametric -----------------------

set.seed(123)
centrality_description(iris, Species, Sepal.Length)

# ----------------------- non-parametric -------------------

set.seed(123)
centrality_description(mtcars, am, wt, type = "n")

# ----------------------- robust ---------------------------

set.seed(123)
centrality_description(ToothGrowth, supp, len, type = "r")

# ----------------------- Bayesian -------------------------

set.seed(123)
centrality_description(sleep, group, extra, type = "b")
```

# Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: <https://github.com/IndrajeetPatil/statsExpressions/issues>
