---
title: "Return value schema"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Return value schema}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label = "setup",
#| message = FALSE,
#| warning = FALSE,
#| include = FALSE,
#| echo = FALSE
source("../setup.R")

pkgs <- c("metaplus", "metafor", "metaBMA")

successfully_loaded <- purrr::map_lgl(pkgs, requireNamespace, quietly = TRUE)
can_evaluate <- all(successfully_loaded)

if (can_evaluate) {
  purrr::walk(pkgs, library, character.only = TRUE)
} else {
  knitr::opts_chunk$set(eval = FALSE)
}
```

This vignette can be cited as:

```{r}
#| label = "citation",
#| echo = FALSE,
#| comment = ""
citation("statsExpressions")
```

## Introduction

Every exported analysis function in `{statsExpressions}` returns a **tidy tibble**
(with the additional class `"statsExpressions"`) rather than a model object. This
means the output is immediately pipe-friendly and ready for downstream use in data
wrangling or graphing pipelines.

Two design principles are worth keeping in mind while reading this article:

1.  **The exact columns depend on the analysis.** All functions follow the same
    tidy/[`{broom}`](https://broom.tidymodels.org/)-style naming conventions, but
    the precise set of columns returned depends on both the *test* and the *type*
    of analysis (`"parametric"`, `"nonparametric"`, `"robust"`, or `"bayes"`). For
    example, Bayesian analyses return Bayes-factor columns (`bf10`, `log_e_bf10`)
    that frequentist analyses do not.

2.  **The `expression` column is the core deliverable.** Regardless of the test,
    almost every function returns a pre-formatted `expression` column intended for
    annotating plots. This is what powers the statistical details you see in
    [`{ggstatsplot}`](https://indrajeetpatil.github.io/ggstatsplot/) plots.

The rest of this article documents the `expression` column, the internal engine
that builds it, and then provides a column-by-column schema for each exported
function.

## The `expression` column

The `expression` column is not a character string. It is a
[plotmath](https://rdrr.io/r/grDevices/plotmath.html) `language` object, which is
why it can render mathematical notation (subscripts, Greek letters, italicized
statistics) when used as a plot label.

```{r}
#| label = "expression-class"
set.seed(123)
res <- two_sample_test(ToothGrowth, supp, len)

# the column is a list of `language` objects, not strings
class(res$expression[[1]])
```

Because it is a `language` object, it can be dropped directly into any
`{ggplot2}` layer that accepts a plotmath expression, such as `labs()`,
`annotate()`, or `ggtitle()`:

```{r}
#| label = "expression-usage",
#| eval = FALSE
library(ggplot2)

ggplot(ToothGrowth, aes(supp, len)) +
  geom_boxplot() +
  labs(subtitle = res$expression[[1]])
```

## The `add_expression_col()` engine

All of the `expression` columns are produced by a single internal engine:
`add_expression_col()`. Understanding it helps clarify why the output looks the
way it does.

The engine works as follows:

-   It consumes a tidy data frame of statistical details -- ideally the output of
    `tidy_model_parameters()`, which standardizes the results from the various
    modeling backends (`{effectsize}`, `{parameters}`, `{BayesFactor}`, etc.) into
    consistent column names.

-   It inspects the available columns to decide which expression *template* to
    use. The key branch points are:

    -   whether the analysis is **Bayesian** (detected via the presence of a
        `bf10` column), and

    -   the number of **degrees-of-freedom parameters** present (`0`, `1`, or
        `2`), which determines whether the statistic is displayed as, e.g.,
        `t = ...`, `t(df) = ...`, or `F(df1, df2) = ...`.

-   It then formats the numeric values (respecting the `digits`, `digits.df`, and
    `digits.df.error` arguments), assembles the plotmath string via
    [`{glue}`](https://glue.tidyverse.org/), converts it to a `language` object,
    and appends it as the `expression` column.

> **Note:** `add_expression_col()` is exported so that its behavior is documented
> and inspectable, but it is **not stable** and is intended for internal use only.
> Its signature and behavior may change without a deprecation cycle. Build
> downstream code against the *output columns* documented here rather than by
> calling this function directly.

## Per-function schema

The examples below use `dplyr::glimpse()` so that every returned column (and its
type) is visible at a glance.

### `one_sample_test()`

```{r}
#| label = "one-sample"
set.seed(123)
one_sample_test(mtcars, wt, test.value = 3) |> dplyr::glimpse()

# Bayesian variant returns Bayes-factor columns instead of a test statistic
set.seed(123)
one_sample_test(mtcars, wt, test.value = 3, type = "bayes") |> dplyr::glimpse()
```

| Column                                                                    | Description                                                          |
| :------------------------------------------------------------------------ | :------------------------------------------------------------------- |
| `mu`                                                                      | the null value the mean is tested against (`test.value`)             |
| `statistic`, `df.error`, `p.value`                                        | test statistic, its degrees of freedom, and *p*-value (non-Bayesian) |
| `effectsize`, `estimate`                                                  | name and value of the effect size (e.g. Hedges' *g*)                 |
| `conf.level`, `conf.low`, `conf.high`, `conf.method`, `conf.distribution` | interval details for the effect size                                 |
| `bf10`, `log_e_bf10`, `prior.*`                                           | Bayes factor, its log, and the prior (Bayesian only)                 |
| `n.obs`                                                                   | number of observations                                               |
| `expression`                                                              | plotmath expression with the test details                            |

### `two_sample_test()`

```{r}
#| label = "two-sample"
set.seed(123)
two_sample_test(ToothGrowth, supp, len) |> dplyr::glimpse()
```

| Column                               | Description                                           |
| :----------------------------------- | :---------------------------------------------------- |
| `parameter1`, `parameter2`           | the outcome and grouping variable names               |
| `mean.parameter1`, `mean.parameter2` | group means (parametric between-subjects)             |
| `statistic`, `df.error`, `p.value`   | test statistic, its degrees of freedom, and *p*-value |
| `effectsize`, `estimate`, `conf.*`   | effect size and its interval                          |
| `n.obs`, `expression`                | sample size and the plotmath expression               |

### `oneway_anova()`

```{r}
#| label = "oneway-anova"
set.seed(123)
oneway_anova(iris, Species, Sepal.Length) |> dplyr::glimpse()
```

| Column                                   | Description                                                  |
| :--------------------------------------- | :----------------------------------------------------------- |
| `statistic`, `df`, `df.error`, `p.value` | *F*-statistic with its two degrees of freedom, and *p*-value |
| `effectsize`, `estimate`, `conf.*`       | effect size (e.g. omega-squared) and its interval            |
| `n.obs`, `expression`                    | sample size and the plotmath expression                      |

For a two-degrees-of-freedom statistic like ANOVA, note that **both** `df` and
`df.error` are present.

### `corr_test()`

```{r}
#| label = "corr-test"
set.seed(123)
corr_test(mtcars, wt, mpg) |> dplyr::glimpse()
```

| Column                                               | Description                                            |
| :--------------------------------------------------- | :----------------------------------------------------- |
| `parameter1`, `parameter2`                           | the two correlated variable names                      |
| `effectsize`, `estimate`                             | name (e.g. Pearson's *r*) and value of the correlation |
| `conf.level`, `conf.low`, `conf.high`, `conf.method` | interval details                                       |
| `statistic`, `df.error`, `p.value`                   | test statistic, degrees of freedom, and *p*-value      |
| `n.obs`, `expression`                                | sample size and the plotmath expression                |

### `contingency_table()`

```{r}
#| label = "contingency"
# two-way association test
set.seed(123)
contingency_table(mtcars, am, vs) |> dplyr::glimpse()

# one-way goodness-of-fit test
set.seed(123)
contingency_table(as.data.frame(HairEyeColor), Eye, counts = Freq) |> dplyr::glimpse()
```

| Column                             | Description                                                  |
| :--------------------------------- | :----------------------------------------------------------- |
| `statistic`, `df`, `p.value`       | chi-squared statistic, its degrees of freedom, and *p*-value |
| `effectsize`, `estimate`, `conf.*` | effect size (e.g. Cramer's *V*) and its interval             |
| `bf10`, `prior.scale`              | Bayes factor and prior (Bayesian only)                       |
| `n.obs`, `expression`              | sample size and the plotmath expression                      |

### `pairwise_comparisons()`

```{r}
#| label = "pairwise-comparisons"
set.seed(123)
pairwise_comparisons(
  mtcars, cyl, wt,
  type = "nonparametric",
  p.adjust.method = "none"
) |> dplyr::glimpse()
```

| Column                    | Description                                             |
| :------------------------ | :------------------------------------------------------ |
| `group1`, `group2`        | the two levels being compared                           |
| `statistic`, `p.value`    | test statistic and *p*-value for the comparison         |
| `p.adjust.method`, `test` | multiple-comparison adjustment method and the test used |
| `expression`              | plotmath expression with the (adjusted) *p*-value       |

Unlike the single-test functions, this returns **one row per pairwise comparison**.

### `pairwise_contingency_table()`

```{r}
#| label = "pairwise-contingency"
set.seed(123)
pairwise_contingency_table(mtcars, cyl, am, p.adjust.method = "holm") |> dplyr::glimpse()
```

| Column                                  | Description                                      |
| :-------------------------------------- | :----------------------------------------------- |
| `group1`, `group2`                      | the two levels being compared                    |
| `p.value`, `p.value.adj`                | unadjusted and multiplicity-adjusted *p*-values  |
| `effectsize`, `estimate`, `conf.*`      | Cramer's *V* effect size and its interval        |
| `p.adjust.method`, `test`, `expression` | adjustment method, test, and plotmath expression |

### `meta_analysis()`

```{r}
#| label = "meta"
# renaming columns to `{statsExpressions}` conventions
data(mag, package = "metaplus")
df <- dplyr::rename(mag, estimate = yi, std.error = sei)

set.seed(123)
meta_analysis(df) |> dplyr::glimpse()
```

| Column                                               | Description                                           |
| :--------------------------------------------------- | :---------------------------------------------------- |
| `term`, `effectsize`                                 | the summary term and effect-size label                |
| `estimate`, `std.error`                              | meta-analytic summary estimate and its standard error |
| `conf.level`, `conf.low`, `conf.high`, `conf.method` | interval details                                      |
| `statistic`, `p.value`                               | test statistic and *p*-value                          |
| `n.obs`, `expression`                                | number of studies and the plotmath expression         |

### `centrality_description()`

```{r}
#| label = "centrality"
set.seed(123)
centrality_description(iris, Species, Sepal.Length) |> dplyr::glimpse()
```

| Column                                                 | Description                                                                        |
| :----------------------------------------------------- | :--------------------------------------------------------------------------------- |
| `<x>` (e.g. `Species`)                                 | the grouping variable                                                              |
| `<y>` (e.g. `Sepal.Length`)                            | the requested centrality measure (mean/median/trimmed mean/MAP)                    |
| `std.dev`, `iqr`, `min`, `max`, `skewness`, `kurtosis` | dispersion and shape statistics                                                    |
| `conf.low`, `conf.high`                                | interval for the centrality estimate                                               |
| `n.obs`, `missing.obs`                                 | count of observations and missing values                                           |
| `expression`                                           | plotmath expression with the centrality estimate                                   |
| `n.expression`                                         | a label combining the group name and its sample size (handy as a facet/axis label) |

This is the only function that returns an `n.expression` column.

## Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: <https://github.com/IndrajeetPatil/statsExpressions/issues>
