---
title: "Test and effect size details"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Test and effect size details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
#| label = "setup",
#| message = FALSE,
#| warning = FALSE,
#| include = FALSE,
#| echo = FALSE
source("../setup.R")

# TODO: check if the pkgdown issue with child document path is resolved
# for now, run `rmarkdown::render("vignettes/web_only/stats_details.Rmd")` from root directory
knitr::opts_chunk$set(
  eval = if (identical(Sys.getenv("IN_PKGDOWN"), "true")) FALSE else TRUE
)
```

This vignette can be cited as:

```{r citation, echo=FALSE, comment = ""}
citation("statsExpressions")
```

## Introduction

Here a go-to summary about statistical test carried out and the returned effect
size for each function is provided. This should be useful if one needs to find
out more information about how an argument is resolved in the underlying package
or if one wishes to browse the source code. So, for example, if you want to know
more about how one-way (between-subjects) ANOVA, you can run
`?stats::oneway.test` in your R console.

Abbreviations used: CI = Confidence Interval

## Summary of functionality

```{r child="../../man/rmd-fragments/functionality.Rmd"}
```

## Summary of tests and effect sizes

Here a go-to summary about statistical test carried out and the returned effect
size for each function is provided. This should be useful if one needs to find
out more information about how an argument is resolved in the underlying package
or if one wishes to browse the source code. So, for example, if you want to know
more about how one-way (between-subjects) ANOVA, you can run
`?stats::oneway.test` in your R console.

### `centrality_description()`

```{r child="../../man/rmd-fragments/centrality_description.Rmd"}
```

### `oneway_anova()`

```{r child="../../man/rmd-fragments/oneway_anova.Rmd"}
```

### `two_sample_test()` 

```{r child="../../man/rmd-fragments/two_sample_test.Rmd"}
```

### `one_sample_test()`

```{r child="../../man/rmd-fragments/one_sample_test.Rmd"}
```

### `corr_test()`

```{r child="../../man/rmd-fragments/corr_test.Rmd"}
```

### `contingency_table()`

```{r child="../../man/rmd-fragments/contingency_table.Rmd"}
```

### `meta_analysis()`

```{r child="../../man/rmd-fragments/meta_analysis.Rmd"}
```

## Effect size interpretation

See `{effectsize}`'s interpretation functions to check different rules/conventions
to interpret effect sizes:

<https://easystats.github.io/effectsize/reference/index.html#section-interpretation>

## References

  - For parametric and non-parametric effect sizes:
    <https://easystats.github.io/effectsize/articles/>

  - For robust effect sizes:
    <https://CRAN.R-project.org/package=WRS2/vignettes/WRS2.pdf>

  - For Bayesian posterior estimates:
    <https://easystats.github.io/bayestestR/articles/bayes_factors.html>

## Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: 
<https://github.com/IndrajeetPatil/statsExpressions/issues>
