---
title: "Dataframe outputs"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    toc_depth: 3
    warning: FALSE
    message: FALSE
vignette: >
  %\VignetteIndexEntry{Dataframe outputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# show me all columns
options(tibble.width = Inf, pillar.bold = TRUE, pillar.subtle_num = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  out.width = "100%",
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

Although the primary focus of this package is to get expressions containing
statistical results, one can also use it to extract dataframes containing these
details. This vignette catalogs dataframe outputs from all functions.

A dataframe will typically contain the following columns:

  - *statistic*: the numeric value of a statistic.

  - *parameter*: the numeric value of a parameter being modeled (often degrees
    of freedom for the test); note that if `no.parameters = 0L` (e.g., for
    non-parametric tests), this column will be irrelevant.

  - *parameter1*, *parameter2* relevant only if the statistic in question has
    two degrees of freedom (e.g., anova).

  - *p.value* the two-sided *p*-value associated with the observed statistic.

  - *estimate*: estimated value of the effect size.

  - *conf.low*: lower bound for effect size estimate.

  - *conf.high*: upper bound for effect size estimate.

For details about the tests and types of effect size estimates they correspond
to, see:
<https://indrajeetpatil.github.io/statsExpressions/articles/stats_details.html>

# One-sample tests

```{r}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
expr_t_onesample(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "parametric",
  output = "dataframe"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
expr_t_onesample(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "nonparametric",
  output = "dataframe"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_t_onesample(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "robust",
  output = "dataframe"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_t_onesample(
  data = ggplot2::msleep,
  x = brainwt,
  test.value = 0.275,
  type = "bayes",
  bf.prior = 0.8,
  output = "dataframe"
)
```

# Two-sample tests

## within-subjects design

```{r}
# setup
library(statsExpressions)

# data
df <- dplyr::filter(bugs_long, condition %in% c("LDLF", "LDHF"))

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
expr_t_parametric(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
expr_t_nonparametric(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_t_robust(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_t_bayes(
  data = df,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)
```

## between-subjects design

```{r}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
# unequal variance
expr_t_parametric(
  data = ToothGrowth,
  x = supp,
  y = len,
  output = "dataframe"
)

# for reproducibility
set.seed(123)
# equal variance
expr_t_parametric(
  data = ToothGrowth,
  x = supp,
  y = len,
  var.equal = TRUE,
  output = "dataframe"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
expr_t_nonparametric(
  data = ToothGrowth,
  x = supp,
  y = len,
  output = "dataframe"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_t_robust(
  data = ToothGrowth,
  x = supp,
  y = len,
  output = "dataframe"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_t_bayes(
  data = ToothGrowth,
  x = supp,
  y = len,
  output = "dataframe"
)
```

# One-way ANOVAs

## within-subjects design

```{r}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
expr_anova_parametric(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
expr_anova_nonparametric(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_anova_robust(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_anova_bayes(
  data = bugs_long,
  x = condition,
  y = desire,
  paired = TRUE,
  subject.id = subject,
  output = "dataframe"
)
```

## between-subjects design

```{r}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
# unequal variance
expr_anova_parametric(
  data = iris,
  x = Species,
  y = Sepal.Length,
  output = "dataframe"
)

# for reproducibility
set.seed(123)
# equal variance
expr_anova_parametric(
  data = iris,
  x = Species,
  y = Sepal.Length,
  var.equal = TRUE,
  output = "dataframe"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
expr_anova_nonparametric(
  data = iris,
  x = Species,
  y = Sepal.Length,
  output = "dataframe"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_anova_robust(
  data = iris,
  x = Species,
  y = Sepal.Length,
  output = "dataframe"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_anova_bayes(
  data = iris,
  x = Species,
  y = Sepal.Length,
  output = "dataframe"
)
```

# Contingency table analyses

```{r}

# for reproducibility
set.seed(123)
library(statsExpressions)

# create data structure
paired_data <-
  structure(
    list(
      response_before =
        structure(
          c(1L, 2L, 1L, 2L),
          .Label = c("no", "yes"),
          class = "factor"
        ),
      response_after = structure(
        c(1L, 1L, 2L, 2L),
        .Label = c("no", "yes"),
        class = "factor"
      ),
      Freq = c(65L, 25L, 5L, 5L)
    ),
    class = "data.frame",
    row.names = c(NA, -4L)
  )

# ------------------------ association tests -----------------------------

# unpaired
set.seed(123)
expr_contingency_tab(
  data = mtcars,
  x = am,
  y = cyl,
  paired = FALSE,
  output = "dataframe"
)

# paired
set.seed(123)
statsExpressions::expr_contingency_tab(
  data = paired_data,
  x = response_before,
  y = response_after,
  paired = TRUE,
  counts = "Freq",
  output = "dataframe"
)


# ------------------------ goodness of fit tests ---------------------------

# with counts
set.seed(123)
expr_contingency_tab(
  data = as.data.frame(HairEyeColor),
  x = Eye,
  counts = Freq,
  output = "dataframe"
)
```

# Correlation analyses

```{r}
# setup
library(statsExpressions)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
expr_corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "parametric",
  output = "dataframe"
)

# ----------------------- non-parametric -----------------------------------

# for reproducibility
set.seed(123)
expr_corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "nonparametric",
  output = "dataframe"
)

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "robust",
  output = "dataframe"
)

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_corr_test(
  data = mtcars,
  x = wt,
  y = mpg,
  type = "bayes",
  output = "dataframe"
)
```

# Meta-analysis

```{r}
# setup
set.seed(123)
library(statsExpressions)
library(metaplus)

# renaming to what `statsExpressions` expects
df <- dplyr::rename(mag, estimate = yi, std.error = sei)

# ----------------------- parametric ---------------------------------------

# for reproducibility
set.seed(123)
expr_meta_random(data = df, output = "dataframe", type = "parametric")

# ----------------------- robust --------------------------------------------

# for reproducibility
set.seed(123)
expr_meta_random(data = df, output = "dataframe", type = "robust")

# ----------------------- Bayes Factor ---------------------------------------

# for reproducibility
set.seed(123)
expr_meta_random(data = df, output = "dataframe", type = "bayes")
```

# Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: <https://github.com/IndrajeetPatil/ggstatsplot/issues>

# Session Information

For details, see- <https://indrajeetpatil.github.io/ggstatsplot/articles/web_only/session_info.html>
