---
title: "Test and effect size details"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Test and effect size details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r citation, echo=FALSE, comment = ""}
citation("statsExpressions")
```

## Introduction

Here a go-to summary about statistical test carried out and the returned effect
size for each function is provided. This should be useful if one needs to find
out more information about how an argument is resolved in the underlying package
or if one wishes to browse the source code. So, for example, if you want to know
more about how one-way (between-subjects) ANOVA, you can run
`?stats::oneway.test` in your R console.

Abbreviations used: CI = Confidence Interval

## Bird's-eye view summary

The table below summarizes all the different types of analyses currently
supported in this package-

Description | Parametric | Non-parametric | Robust | Bayesian
------------------ | ---- | ----- | ----| ----- 
Between group/condition comparisons | ✅ | ✅ | ✅ | ✅
Within group/condition comparisons | ✅ | ✅ | ✅ | ✅
Distribution of a numeric variable | ✅ | ✅ | ✅ | ✅
Correlation between two variables | ✅ | ✅ | ✅ | ✅
Association between categorical variables | ✅ | ✅ | ❌ | ✅
Equal proportions for categorical variable levels | ✅ | ✅ | ❌ | ✅
Random-effects meta-analysis | ✅ | ❌ | ✅ | ✅

Summary of Bayesian analysis

Analysis | Hypothesis testing | Estimation
------------------ | ---------- | ---------
(one/two-sample) *t*-test | ✅ | ✅
one-way ANOVA | ✅ | ✅
correlation | ✅ | ✅
(one/two-way) contingency table | ✅ | ✅
random-effects meta-analysis | ✅ | ✅

## Summary of tests and effect sizes

Here a go-to summary about statistical test carried out and the returned effect
size for each function is provided. This should be useful if one needs to find
out more information about how an argument is resolved in the underlying package
or if one wishes to browse the source code. So, for example, if you want to know
more about how one-way (between-subjects) ANOVA, you can run
`?stats::oneway.test` in your R console.

### `centrality_description`

Type | Measure | Function used
----------- | --------- | ------------------ 
Parametric | mean | `parameters::describe_distribution`
Non-parametric | median | `parameters::describe_distribution`
Robust | trimmed mean | `parameters::describe_distribution`
Bayesian | MAP (maximum *a posteriori* probability) estimate | `parameters::describe_distribution`

### `two_sample_test` + `oneway_anova`

No. of groups: `2` => `two_sample_test`<br>
No. of groups: `> 2` => `oneway_anova`

#### between-subjects

**Hypothesis testing**

Type | No. of groups | Test | Function used
----------- | --- | ------------------------- | -----
Parametric | > 2 | Fisher's or Welch's one-way ANOVA | `stats::oneway.test`
Non-parametric | > 2 | Kruskal–Wallis one-way ANOVA | `stats::kruskal.test`
Robust | > 2 | Heteroscedastic one-way ANOVA for trimmed means | `WRS2::t1way`
Bayes Factor | > 2 | Fisher's ANOVA | `BayesFactor::anovaBF`
Parametric | 2 | Student's or Welch's *t*-test | `stats::t.test`
Non-parametric | 2 | Mann–Whitney *U* test | `stats::wilcox.test`
Robust | 2 |  Yuen's test for trimmed means | `WRS2::yuen`
Bayesian | 2 | Student's *t*-test | `BayesFactor::ttestBF`

**Effect size estimation**

Type | No. of groups | Effect size | CI? | Function used
----------- | --- | ------------------------- | --- | -----
Parametric | > 2 | $\eta_{p}^2$, $\omega_{p}^2$ | ✅ | `effectsize::omega_squared`, `effectsize::eta_squared`
Non-parametric | > 2 | $\epsilon_{ordinal}^2$ | ✅ | `effectsize::rank_epsilon_squared`
Robust | > 2 | $\xi$ (Explanatory measure of effect size) | ✅ | `WRS2::t1way`
Bayes Factor | > 2 | $R_{Bayesian}^2$ | ✅ | `performance::r2_bayes`
Parametric | 2 | Cohen's *d*, Hedge's *g* | ✅ | `effectsize::cohens_d`, `effectsize::hedges_g`
Non-parametric | 2 | *r* (rank-biserial correlation) | ✅ | `effectsize::rank_biserial`
Robust | 2 |  $\delta_{R}^{AKP}$ (Algina-Keselman-Penfield robust standardized difference) | ✅ | `WRS2::akp.effect`
Bayesian | 2 | $\delta_{posterior}$ | ✅ | `bayestestR::describe_posterior`

#### within-subjects

**Hypothesis testing**

Type | No. of groups | Test | Function used
----------- | --- | ------------------------- | -----
Parametric | > 2 | One-way repeated measures ANOVA | `afex::aov_ez`
Non-parametric | > 2 | Friedman rank sum test | `stats::friedman.test`
Robust | > 2 | Heteroscedastic one-way repeated measures ANOVA for trimmed means | `WRS2::rmanova`
Bayes Factor | > 2 | One-way repeated measures ANOVA | `BayesFactor::anovaBF`
Parametric | 2 | Student's *t*-test | `stats::t.test`
Non-parametric | 2 | Wilcoxon signed-rank test | `stats::wilcox.test`
Robust | 2 | Yuen's test on trimmed means for dependent samples | `WRS2::yuend`
Bayesian | 2 | Student's *t*-test | `BayesFactor::ttestBF`

**Effect size estimation**

Type | No. of groups | Effect size | CI? | Function used
----------- | --- | ------------------------- | --- | -----
Parametric | > 2 | $\eta_{p}^2$, $\omega_{p}^2$ | ✅ | `effectsize::omega_squared`, `effectsize::eta_squared`
Non-parametric | > 2 | $W_{Kendall}$ (Kendall's coefficient of concordance) | ✅ | `effectsize::kendalls_w`
Robust | > 2 | $\delta_{R-avg}^{AKP}$ (Algina-Keselman-Penfield robust standardized difference average) | ✅ | `WRS2::wmcpAKP`
Bayes Factor | > 2 | $R_{Bayesian}^2$ | ✅ | `performance::r2_bayes`
Parametric | 2 | Cohen's *d*, Hedge's *g* | ✅ | `effectsize::cohens_d`, `effectsize::hedges_g`
Non-parametric | 2 | *r* (rank-biserial correlation) | ✅ | `effectsize::rank_biserial`
Robust | 2 |  $\delta_{R}^{AKP}$ (Algina-Keselman-Penfield robust standardized difference) | ✅ | `WRS2::wmcpAKP`
Bayesian | 2 | $\delta_{posterior}$ | ✅ | `bayestestR::describe_posterior`

### `one_sample_test`

**Hypothesis testing**

Type | Test | Function used
------------------ | ------------------------- | -----
Parametric | One-sample Student's *t*-test | `stats::t.test`
Non-parametric | One-sample Wilcoxon test | `stats::wilcox.test`
Robust | Bootstrap-*t* method for one-sample test | `WRS2::trimcibt`
Bayesian | One-sample Student's *t*-test | `BayesFactor::ttestBF`

**Effect size estimation**

Type | Effect size | CI? | Function used
------------ | ----------------------- | --- | -----
Parametric | Cohen's *d*, Hedge's *g* | ✅ | `effectsize::cohens_d`, `effectsize::hedges_g`
Non-parametric | *r* (rank-biserial correlation) | ✅ | `effectsize::rank_biserial`
Robust | trimmed mean | ✅ | `trimcibt` (custom)
Bayes Factor | $\delta_{posterior}$ | ✅ | `bayestestR::describe_posterior`

### `corr_test`

```{r child="../man/rmd-fragments/corr_test.Rmd"}
```

### `contingency_table`

#### two-way table

**Hypothesis testing**

Type | Design | Test | Function used
----------- | ----- | ------------------------- | -----
Parametric/Non-parametric | Unpaired | Pearson's $\chi^2$ test | `stats::chisq.test`
Bayesian | Unpaired | Bayesian Pearson's $\chi^2$ test | `BayesFactor::contingencyTableBF`
Parametric/Non-parametric | Paired  | McNemar's $\chi^2$ test | `stats::mcnemar.test`
Bayesian | Paired  | ❌ | ❌

**Effect size estimation**

Type | Design | Effect size | CI? | Function used
----------- | ----- | ------------------------- | --- | -----
Parametric/Non-parametric | Unpaired | Cramer's $V$ | ✅ | `effectsize::cramers_v`
Bayesian | Unpaired | Cramer's $V$ | ✅ | `effectsize::cramers_v`
Parametric/Non-parametric | Paired | Cohen's $g$ | ✅ | `effectsize::cohens_g`
Bayesian | Paired | ❌ | ❌ | ❌

#### one-way table

**Hypothesis testing**

Type | Test | Function used
----------- | ------------------------- | -----
Parametric/Non-parametric | Goodness of fit $\chi^2$ test | `stats::chisq.test`
Bayesian | Bayesian Goodness of fit $\chi^2$ test | (custom)

**Effect size estimation**

Type | Effect size | CI? | Function used
----------- | ------------------------- | --- | -----
Parametric/Non-parametric | Pearson's $C$ | ✅ | `effectsize::pearsons_c`
Bayesian | ❌ | ❌ | ❌

### `meta_analysis`

**Hypothesis testing** and **Effect size estimation**

Type | Test | Effect size | CI? | Function used
----------- | -------------------- | -------- | ---  | -----
Parametric | Meta-analysis via random-effects models | $\beta$ | ✅ | `metafor::metafor`
Robust | Meta-analysis via robust random-effects models | $\beta$ | ✅ | `metaplus::metaplus`
Bayes | Meta-analysis via Bayesian random-effects models | $\beta$ | ✅ | `metaBMA::meta_random`

## Effect size interpretation

See `effectsize`'s interpretation functions to check different rules/conventions
to interpret effect sizes:

<https://easystats.github.io/effectsize/reference/index.html#section-interpretation>

## Dataframe as output

Although the primary focus of this package is to get expressions containing
statistical results, one can also use it to extract dataframes containing these
details. 

For a more detailed summary of these dataframe:
<https://indrajeetpatil.github.io/statsExpressions//articles/web_only/dataframe_outputs.html>

## References

  - For parametric and non-parametric effect sizes:
    <https://easystats.github.io/effectsize/articles/simple_htests.html>

  - For robust effect sizes:
    <https://CRAN.R-project.org/package=WRS2/vignettes/WRS2.pdf>

  - For Bayesian posterior estimates:
    <https://easystats.github.io/bayestestR/articles/bayes_factors.html>

## Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: 
<https://github.com/IndrajeetPatil/statsExpressions/issues>
