---
title: "Test and effect size details"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Test and effect size details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r citation, echo=FALSE, comment = ""}
citation("statsExpressions")
```

## Introduction

Here a go-to summary about statistical test carried out and the returned effect
size for each function is provided. This should be useful if one needs to find
out more information about how an argument is resolved in the underlying package
or if one wishes to browse the source code. So, for example, if you want to know
more about how one-way (between-subjects) ANOVA, you can run
`?stats::oneway.test` in your R console.

Abbreviations used: CI = Confidence Interval

## Bird's-eye view summary

The table below summarizes all the different types of analyses currently
supported in this package-

Description | Parametric | Non-parametric | Robust | Bayesian
------------------ | ---- | ----- | ----| ----- 
Between group/condition comparisons | Yes | Yes | Yes | Yes
Within group/condition comparisons | Yes | Yes | Yes | Yes
Distribution of a numeric variable | Yes | Yes | Yes | Yes
Correlation between two variables | Yes | Yes | Yes | Yes
Association between categorical variables | Yes | Yes | No | Yes
Equal proportions for categorical variable levels | Yes | Yes | No | Yes
Random-effects meta-analysis | Yes | No | Yes | Yes

Summary of Bayesian analysis

Analysis | Hypothesis testing | Estimation
------------------ | ---------- | ---------
(one/two-sample) *t*-test | Yes | Yes
one-way ANOVA | Yes | Yes
correlation | Yes | Yes
(one/two-way) contingency table | Yes | Yes
random-effects meta-analysis | Yes | Yes

## Summary of tests and effect sizes

Here a go-to summary about statistical test carried out and the returned effect
size for each function is provided. This should be useful if one needs to find
out more information about how an argument is resolved in the underlying package
or if one wishes to browse the source code. So, for example, if you want to know
more about how one-way (between-subjects) ANOVA, you can run
`?stats::oneway.test` in your R console.

### `centrality_description`

Type | Measure | Function used
----------- | --------- | ------------------ 
Parametric | mean | `parameters::describe_distribution`
Non-parametric | median | `parameters::describe_distribution`
Robust | trimmed mean | `parameters::describe_distribution`
Bayesian | MAP (maximum *a posteriori* probability) estimate | `parameters::describe_distribution`

### `oneway_anova`

```{r child="../man/rmd-fragments/oneway_anova.Rmd"}
```

### `two_sample_test` 

```{r child="../man/rmd-fragments/two_sample_test.Rmd"}
```

### `one_sample_test`

```{r child="../man/rmd-fragments/one_sample_test.Rmd"}
```

### `corr_test`

```{r child="../man/rmd-fragments/corr_test.Rmd"}
```

### `contingency_table`

#### two-way table

**Hypothesis testing**

Type | Design | Test | Function used
----------- | ----- | ------------------------- | -----
Parametric/Non-parametric | Unpaired | Pearson's $\chi^2$ test | `stats::chisq.test`
Bayesian | Unpaired | Bayesian Pearson's $\chi^2$ test | `BayesFactor::contingencyTableBF`
Parametric/Non-parametric | Paired  | McNemar's $\chi^2$ test | `stats::mcnemar.test`
Bayesian | Paired  | No | ❌

**Effect size estimation**

Type | Design | Effect size | CI available? | Function used
----------- | ----- | ------------------------- | --- | -----
Parametric/Non-parametric | Unpaired | Cramer's $V$ | Yes | `effectsize::cramers_v`
Bayesian | Unpaired | Cramer's $V$ | Yes | `effectsize::cramers_v`
Parametric/Non-parametric | Paired | Cohen's $g$ | Yes | `effectsize::cohens_g`
Bayesian | Paired | No | No | ❌

#### one-way table

**Hypothesis testing**

Type | Test | Function used
----------- | ------------------------- | -----
Parametric/Non-parametric | Goodness of fit $\chi^2$ test | `stats::chisq.test`
Bayesian | Bayesian Goodness of fit $\chi^2$ test | (custom)

**Effect size estimation**

Type | Effect size | CI available? | Function used
----------- | ------------------------- | --- | -----
Parametric/Non-parametric | Pearson's $C$ | Yes | `effectsize::pearsons_c`
Bayesian | No | No | ❌

### `meta_analysis`

**Hypothesis testing** and **Effect size estimation**

Type | Test | Effect size | CI available? | Function used
----------- | -------------------- | -------- | ---  | -----
Parametric | Meta-analysis via random-effects models | $\beta$ | Yes | `metafor::metafor`
Robust | Meta-analysis via robust random-effects models | $\beta$ | Yes | `metaplus::metaplus`
Bayes | Meta-analysis via Bayesian random-effects models | $\beta$ | Yes | `metaBMA::meta_random`

## Effect size interpretation

See `effectsize`'s interpretation functions to check different rules/conventions
to interpret effect sizes:

<https://easystats.github.io/effectsize/reference/index.html#section-interpretation>

## Dataframe as output

Although the primary focus of this package is to get expressions containing
statistical results, one can also use it to extract dataframes containing these
details. 

For a more detailed summary of these dataframe:
<https://indrajeetpatil.github.io/statsExpressions//articles/web_only/dataframe_outputs.html>

## References

  - For parametric and non-parametric effect sizes:
    <https://easystats.github.io/effectsize/articles/simple_htests.html>

  - For robust effect sizes:
    <https://CRAN.R-project.org/package=WRS2/vignettes/WRS2.pdf>

  - For Bayesian posterior estimates:
    <https://easystats.github.io/bayestestR/articles/bayes_factors.html>

## Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: 
<https://github.com/IndrajeetPatil/statsExpressions/issues>
