---
title: "Gallery of examples"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    toc_depth: 3
    warning: FALSE
    message: FALSE
    eval: FALSE
vignette: >
  %\VignetteIndexEntry{Gallery of examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label = "setup",
#| include = FALSE
# show me all columns
options(
  tibble.width = Inf,
  pillar.bold = TRUE,
  pillar.neg = TRUE,
  pillar.subtle_num = TRUE,
  pillar.min_chars = Inf
)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  out.width = "100%",
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  eval = FALSE
)
```

# Introduction

`statsExpressions` dataframes contain *expression*, meaning **a pre-formatted in-text statistical result**.

These expressions can be used to display statistical analysis results for any
`ggplot` plot, which means it can also be used with any of the `ggplot2`
extension packages. This vignette provides a gallery of such examples.

# Examples

## `ggforce`

```{r}
#| label = "ggforce1"
# setup
set.seed(123)
library(ggplot2)
library(ggforce)
library(statsExpressions)
# plot with subtitle
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  geom_violin() +
  geom_sina() +
  labs(
    title = "Fisher's one-way ANOVA",
    subtitle = oneway_anova(iris, Species, Sepal.Length, var.equal = TRUE)$expression[[1]]
  )
```

## `ggExtra`

```{r}
#| label = "ggExtra1"
# setup
set.seed(123)
library(ggplot2)
library(ggExtra)
library(statsExpressions)
# basic plot
p <-
  ggplot(mtcars, aes(mpg, wt)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Pearson's correlation coefficient",
    subtitle = corr_test(mtcars, mpg, wt, type = "parametric")$expression[[1]]
  )
# add
ggMarginal(p, type = "histogram", xparams = list(binwidth = 1, fill = "orange"))
```

## `ggpubr`

```{r}
#| label = "ggpubr1"
set.seed(123)
library(ggpubr)
library(ggplot2)

# plot
ggboxplot(
  ToothGrowth,
  x = "dose",
  y = "len",
  color = "dose",
  palette = c("#00AFBB", "#E7B800", "#FC4E07"),
  add = "jitter",
  shape = "dose"
) + # adding results from stats analysis using `statsExpressions`
  labs(
    title = "Kruskall-Wallis test",
    subtitle = oneway_anova(ToothGrowth, dose, len, type = "np")$expression[[1]]
  )
```

```{r}
#| label = "ggpubr2"
# setup
set.seed(123)
library(ggplot2)
library(ggpubr)
library(statsExpressions)

# basic plot
gghistogram(
  data.frame(
    sex = factor(rep(c("F", "M"), each = 200)),
    weight = c(rnorm(200, 55), rnorm(200, 58))
  ),
  x = "weight",
  add = "mean",
  rug = TRUE,
  fill = "sex",
  palette = c("#00AFBB", "#E7B800"),
  add_density = TRUE
) + # displaying stats results
  labs(
    title = "Yuen's two-sample test for trimmed means",
    subtitle = two_sample_test(
      data = data.frame(
        sex = factor(rep(c("F", "M"), each = 200)),
        weight = c(rnorm(200, 55), rnorm(200, 58))
      ),
      x = sex,
      y = weight,
      type = "robust"
    )$expression[[1]]
  )
```

## `ggiraphExtra`

```{r}
#| label = "ggiraphExtra1"
# setup
set.seed(123)
library(ggplot2)
library(ggiraphExtra)
library(gcookbook)
library(statsExpressions)

# plot
ggDot(heightweight, aes(sex, heightIn, fill = sex),
  boxfill = "white",
  binwidth = 0.4
) +
  labs(
    title = "Mann-Whitney test",
    subtitle = two_sample_test(heightweight, sex, heightIn, type = "nonparametric")$expression[[1]]
  )
```

```{r}
#| label = "ggiraphExtra2"
# setup
set.seed(123)
library(moonBook)
library(ggiraphExtra)
library(statsExpressions)

# plot
ggSpine(
  data = acs,
  aes(x = Dx, fill = smoking),
  addlabel = TRUE,
  interactive = FALSE
) +
  labs(
    x = "diagnosis",
    title = "Pearson's chi-squared contingency table test for counts",
    subtitle = contingency_table(acs, Dx, smoking, paired = FALSE)$expression[[1]]
  )
```

# Suggestions

If you find any bugs or have any suggestions/remarks, please file an issue on GitHub: <https://github.com/IndrajeetPatil/ggstatsplot/issues>
